A study on student performance, engagement and experience with Kaggle-in-class data challenges

### 2021-01-07 paper-kaggle.Rmd
done with changes and response
done with all changes for 4th submission

---
title: Student performance, engagement and experience when teaching includes Kaggle-in-class data challenges

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|r|r|}\hline
Question Set & Median difference & Permutation $p$-value \\\hline
Classification & 0.250 & 0.03\\
Regression & 0.104 & 0.00\\\hline
\end{tabular}
\caption{Comparison of median difference in performance by competition group, for CSDM students, using permutation tests. Both sets of medians are significantly different, indicating improved scores for questions on the topic related to the Kaggle competition.}
\end{center}
\label{tab:Melb_Perm}
\end{table}

Table 3 shows the results of permutation testing of median difference between the groups. Generally the results support the competition improved performance. Students who participated in the kaggle challenge for classification scored higher than those that did the regression competition, on the classification problem. Using a permutation test, this corresponds to a significant difference in medians. Similarly the results show that students who did the regression challenge, performed better on these exam questions.

@@

Based on the median, the students who participated in the kaggle challenge scored `r round(dif, 2)` higher than those that didn't, a median of `r round(median(scores$perf[scores$class=="Yes (ST-PG)"]), 2)` in comparison to `r round(median(scores$perf[scores$class=="No (ST-UG)"]), 2)`. Using a permutation test, this corresponds to a significant difference in medians, with $p$-value of `r round(length(pdif[pdif >= dif])/1000, 2)`.

@@

The number of submissions that a student made, may be an indicator of performance on the exam questions related to the competition.  A student who is more engaged in the competition may learn more about the material, and consequently perform better on the exam. Figure \ref{fig:numsubmition} (top row) shows performance on the classification and regression questions, respectively, against their frequency of prediction submissions for the three student groups (CSDM classification and regression, ST-PG regression) competitions. The relationship is weak in all groups, and this mirrors insignificant results from a linear model fit to both subsets. On the other hand, the predictive accuracy improved with the number of submissions for the regression competitions. There appears to be some nonlinearity present in these plots, suggesting reduced returns. That is reasonable to expect. Also, some students strategically make very poor initial predictions, to get a baseline on error equivalent to guessing.

@@

@@

In awarding course points to student effort, we typically align it to performance. Participant ranks based on their performance on the private part of the test data are recorded. Performance scores that are pretty close to each other should be given the same rank, reflecting that there may not be a significant difference between them. The best gets perhaps, 5 points, then a half a point drop until about 2.5 points, so that the worst performing students still get 50% for the task. Kaggle does not allow you to download participants email addresses. All you see is their kaggle name. Record the student names in kaggle to match with your class records.

@@

This paper has described an experiment to examine the effectiveness of data competitions on student learning, using Kaggle InClass as the vehicle for conducting the competition. The primary finding, is that participating in a data challenge competition,  produces a statistically significant improvement in the learning of the topic, although the effect size is small.  Secondarily, the competitions enhanced interest and engagement in the course.

>>>>

---
title: A study on student performance, engagement and experience with Kaggle-in-class data challenges

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|r|r|}\hline
Question Set & Median difference & Permutation $p$-value \\\hline
Classification & 0.250 & 0.03\\
Regression & 0.104 & 0.00\\\hline
\end{tabular}
\caption{Comparison of median difference in performance by competition group, for CSDM students, using permutation tests. Both sets of medians are discernible different, indicating improved scores for questions on the topic related to the Kaggle competition.}
\end{center}
\label{tab:Melb_Perm}
\end{table}

Table 3 shows the results of permutation testing of median difference between the groups. Generally the results support the competition improved performance. Students who participated in the kaggle challenge for classification scored higher than those that did the regression competition, on the classification problem. Using a permutation test, this corresponds to a discernible difference in medians. Similarly the results show that students who did the regression challenge, performed better on these exam questions.

@@

Based on the median, the students who participated in the kaggle challenge scored `r round(dif, 2)` higher than those that didn't, a median of `r round(median(scores$perf[scores$class=="Yes (ST-PG)"]), 2)` in comparison to `r round(median(scores$perf[scores$class=="No (ST-UG)"]), 2)`. Using a permutation test, this corresponds to a discernible difference in medians, with $p$-value of `r round(length(pdif[pdif >= dif])/1000, 2)`.

@@

The number of submissions that a student made, may be an indicator of performance on the exam questions related to the competition.  A student who is more engaged in the competition may learn more about the material, and consequently perform better on the exam. Figure \ref{fig:numsubmition} (top row) shows performance on the classification and regression questions, respectively, against their frequency of prediction submissions for the three student groups (CSDM classification and regression, ST-PG regression) competitions. The relationship is weak in all groups, and this mirrors indiscernible results from a linear model fit to both subsets. On the other hand, the predictive accuracy improved with the number of submissions for the regression competitions. There appears to be some nonlinearity present in these plots, suggesting reduced returns. That is reasonable to expect. Also, some students strategically make very poor initial predictions, to get a baseline on error equivalent to guessing.

@@

Students mostly agree that taking part in data competition improved their understanding of the covered material (Q3) and their skills to apply the covered material to real problems (Q5). Surprisingly, they didn't perceive the Kaggle challenge to be related to exam performance (Q4). Perhaps the link between the two should be emphasised by the teaching stuff when the competition is presented to students. We believe that by increasing students' awareness of the linkage between the knowledge obtained from the data competition, better understanding of the material and better marks, further increases students' engagement with the competition. This, in turn, will positively reflect on their exam scores.

@@

In awarding course points to student effort, we typically align it to performance. Participant ranks based on their performance on the private part of the test data are recorded. Performance scores that are pretty close to each other should be given the same rank, reflecting that there may not be a discernible difference between them. The best gets perhaps, 5 points, then a half a point drop until about 2.5 points, so that the worst performing students still get 50% for the task. Kaggle does not allow you to download participants email addresses. All you see is their kaggle name. Record the student names in kaggle to match with your class records.

@@

This paper has described an experiment to examine the effectiveness of data competitions on student learning, using Kaggle InClass as the vehicle for conducting the competition. The primary finding, is that participating in a data challenge competition, produces a statistically discernible improvement in the learning of the topic, although the effect size is small.  Secondarily, the competitions enhanced interest and engagement in the course.
